---
title: "Data Source"
subtitle: "Federal Reserve Economic Data"
date: "2023-12-3"
date-modified: today
date-format: long
format: 
  html:
    smooth-scroll: true
    toc: true
    code-fold: true
    code-tools: true
    embed-resources: true
    lightbox: true
bibliography: citation.bib
execute: 
  echo: false
  warning: false
---

***
### Federal Reserve Economic Data

FRED is an online database maintained by the Research Division of the Federal Reserve Bank of St. Louis. It holds more than 800,000 economic time series from over 100 sources, including the Federal Reserve Board, the Bureau of Labor Statistics, the Bureau of Economic Analysis, and the U.S. Treasury. Access is free. Every series has a fixed identifier, a stated frequency and unit, and a record of past revisions, which makes the data straightforward to cite and to reproduce. All series in this project come from @fred.

Each series was downloaded as a spreadsheet with two columns: `observation_date` and the observed value, with the value column named by its FRED identifier. Downloads were made in late 2023, so most series end between August and October of that year.

### Series

The table lists every series used in the project. The FRED identifier links to the series page, which documents the source agency, the release schedule, and any seasonal adjustment.

| Series | FRED ID | Frequency | Units | Range | Used in |
|:---|:---|:---|:---|:---|:---|
| Federal Funds Effective Rate | [FEDFUNDS](https://fred.stlouisfed.org/series/FEDFUNDS) | Monthly | Percent | 1954–2023 | ARIMAX, ARCH/GARCH, Deep Learning, EDA, Visualization |
| 3-Month Treasury Bill minus Fed Funds | [T3MFF](https://fred.stlouisfed.org/series/T3MFF) | Daily | Percentage points | 1982–2023 | ARMA, ARIMAX, EDA, Visualization |
| 6-Month Treasury Bill minus Fed Funds | [T6MFF](https://fred.stlouisfed.org/series/T6MFF) | Daily | Percentage points | 1982–2023 | ARMA, ARIMAX, EDA, Visualization |
| 1-Year Treasury minus Fed Funds | [T1YFF](https://fred.stlouisfed.org/series/T1YFF) | Daily | Percentage points | 1962–2023 | ARIMAX, EDA, Visualization |
| 5-Year Treasury minus Fed Funds | [T5YFF](https://fred.stlouisfed.org/series/T5YFF) | Daily | Percentage points | 1962–2023 | ARIMAX |
| 10-Year Treasury, Constant Maturity | [DGS10](https://fred.stlouisfed.org/series/DGS10) | Daily | Percent | 1962–2023 | ARIMAX, EDA, Visualization |
| 20-Year Treasury, Constant Maturity | [DGS20](https://fred.stlouisfed.org/series/DGS20) | Daily | Percent | 1962–2023 | EDA, Visualization |
| 30-Year Treasury, Constant Maturity | [DGS30](https://fred.stlouisfed.org/series/DGS30) | Daily | Percent | 1977–2023 | ARIMAX, EDA, Visualization |
| 30-Year Fixed Rate Mortgage Average | [MORTGAGE30US](https://fred.stlouisfed.org/series/MORTGAGE30US) | Weekly | Percent | 1971–2023 | ARIMAX, EDA, Visualization |
| Consumer Price Index, All Urban Consumers | [CPIAUCSL](https://fred.stlouisfed.org/series/CPIAUCSL) | Monthly | Index, 1982–84 = 100 | 1947–2023 | ARMA, ARIMAX, EDA, Visualization |
| Real Gross Domestic Product | [GDPC1](https://fred.stlouisfed.org/series/GDPC1) | Quarterly | \$ billions, chained 2017 | 1947–2023 | ARMA, ARIMAX, EDA, Visualization |
| Unemployment Rate | [UNRATE](https://fred.stlouisfed.org/series/UNRATE) | Monthly | Percent | 1948–2023 | ARIMAX, EDA, Visualization |
| Consumer Sentiment, University of Michigan | [UMCSENT](https://fred.stlouisfed.org/series/UMCSENT) | Monthly | Index, 1966 Q1 = 100 | 1952–2023 | Deep Learning |
| Nominal Broad U.S. Dollar Index | [DTWEXBGS](https://fred.stlouisfed.org/series/DTWEXBGS) | Daily | Index, Jan 2006 = 100 | 2006–2023 | ARCH/GARCH, EDA, Visualization |
| S&P 500 | [SP500](https://fred.stlouisfed.org/series/SP500) | Daily | Index level | 2013–2023 | ARCH/GARCH |

: {.striped .hover tbl-colwidths="[30,13,9,17,10,21]"}

### Spreads, Not Yields

Four of the identifiers end in `FF`. These are not Treasury yields. Each is the yield at that maturity minus the effective Federal Funds Rate on the same day, in % points. A value of 0.50 for `T3MFF` means the 3-month bill yielded half a point more than the policy rate. The spreads measure how far market rates at the short end sit from the Fed's target, which is the relationship this project is built around. They are the inputs to the ARMA section and to the short-term visualizations, and, with the policy rate added back, they supply the short-maturity yields in the ARIMAX section's yield-curve VAR. The `DGS` series are the yields themselves, at constant maturity, and cover the long end of the curve.

### Frequency

The series arrive at four frequencies. Treasury yields, the spreads, and the dollar index are daily, on Treasury market trading days. The mortgage rate is weekly, published on Thursdays. The Federal Funds Rate, CPI, unemployment, and consumer sentiment are monthly. Real GDP is quarterly. Where a model needs a single frequency, the analysis section states how the series were aligned.

### Merged Files

Two files combine several series on a common date index. `sr_t-bill-rates` holds the three short-maturity spreads on the daily Treasury calendar from January 1982, with no gaps. `t-bill-vs-econ-index` places nine series on the same daily calendar from January 1962. The lower-frequency series in that file are not filled forward: CPI, the Federal Funds Rate, unemployment, and GDP appear only on their release dates and are empty on every other row. The previews below show the structure.

```{r}
suppressMessages(library(readxl))
library(knitr)
options(knitr.kable.NA = "—")

sr <- read_excel("data/data/sr_t-bill-rates.xlsx")
sr$observation_date <- as.Date(sr$observation_date)

mx <- read_excel("data/data/t-bill-vs-econ-index.xlsx",
                 col_types = c("date", rep("numeric", 9)))
mx$observation_date <- as.Date(mx$observation_date)
```

::: panel-tabset
## Short-Term Spreads

`sr_t-bill-rates.xlsx` — 10,420 daily rows, 1982–2023.

```{r}
kable(head(sr, 8),
      col.names = c("Date", "T3MFF", "T6MFF", "T1YFF"),
      align = "lrrr", digits = 2)
```

## Yields and Economic Indicators

`t-bill-vs-econ-index.xlsx` — 15,974 daily rows, 1962–2023. Shown from 1 August 2023, a monthly release date.

```{r}
win <- mx[mx$observation_date >= as.Date("2023-08-01"), ][1:8, ]
kable(win,
      col.names = c("Date", names(win)[-1]),
      align = c("l", rep("r", 9)), digits = 2)
```
:::
